Push-Pull Signal-Function Functional Reactive Programming

### 2012-11-29 Docs/IFL2012/implementation.tex
implementation section

The design of signal functions specifies a family of types for the inputs and
outputs of signal functions. Signal functions are not functions in the purest
sense, however. They are not mappings from a single instance of their input
type to a single instance of their output type. They must be implemented with
respect to the temporal semantics of their inputs and outputs.

We therefore start by creating a set of concrete datatypes for the inputs and
outputs of signal functions. These datatypes will be parameterized by the input
and output types of the signal function, and will not be exposed to the user of
the library. Rather, they will specify how data is represented during the
temporal evaluation of signal functions. We then implement signal functions
as records of functions from these concrete types to these concrete types paired
with new signal functions. The evaluation system for signal functions then
maintains this record for the signal function it is evaluating, calls the
correct function for the current input, actuates the outputs (using supplied
functions from the outputs to monadic actions) and replaces the record with the
newly supplied signal function record.

In Section~\ref{section:System_Design_and_Interface} we presented signal vectors
as a set of types. In order to be completely principled, we should isolate these
types into their own {\em kind} (a sort of type of types); however, the Haskell
extension for this was far from stable at the time this system was created.

Using the Glasgow Haskell Compiler's extension for Generalized Algebraic
Datatypes~\cite{Cheney2003,Xi2003,PeytonJones2006}, we specify three concrete
datatypes which are parameterized over signal vectors, and represent
information about one temporal point in a signal function's evaluation.

@@

The next representation contains a constructor for {\tt SVEvent} nodes, carrying
a value typed by the type parameter of {\tt SVEvent}, and two constructors, each
of which carries a representation for one child of a {\tt SVAppend} node (the
left or right), and is parametric in the other. Thus, an instance of this
representation is a path from the root of a signal vector to an event node,
carrying exactly one value for the event.

@@

This is the type and framework for our signal function implementation.
Combinators are implemented as functions which return these types with specific
implementations of the continuations. Space precludes a full discussion of the
implemenation of each combinator, but some of the examples will be discussed.

@@

Time dependence is introduced by the {\tt time}, {\tt delay}, and {\tt integrate}
combinators. The time combinator simply sums the time updates and provides the
sum as a signal output. The {\tt delay} combinator keeps a table of events
which have come in, along with their schedule occurrence time, and produces
them as output when time advances far enough. The integrate combinator performs
rectangle-rule integration on signal samples with respect to time.

>>>>

The design of signal functions specifies a family of types for the inputs and
outputs of signal functions. Signal functions are not functions in the purest
sense, however. They are not mappings from a single instance of their input
type to a single instance of their output type. They must be implemented with
respect to the temporal semantics of their inputs and outputs.
\rn{What does that mean?  ``implement with respect to the temporal
  semantics'' seems like an underspecified relation.  I'm also not
  sure about the use of ``purest'' -- isn't this a semantics
  vs. runtime/implementation issue?}

We therefore start by creating a set of concrete datatypes for the inputs and
outputs of signal functions. These datatypes will be parameterized by the input
and output types of the signal function, and will not be exposed to the user of
the library. Rather, they will specify how data is represented during the
temporal evaluation of signal functions. 
\rn{Maybe ``temporal evaluation'' could be more concrete... this means
  repeated running sampling computation, right?  Maybe there should be a term
  for each sampling interval, epoch, whatever?}
We then implement signal functions
as records of functions from these concrete types \textred{to these concrete types} paired
with new signal functions. The evaluation system for signal functions then
maintains this record for the signal function it is evaluating, calls the
correct function for the current input, actuates the outputs (using supplied
functions from the outputs to monadic actions) and replaces the record with the
newly supplied signal function record.

In Section~\ref{section:System_Design_and_Interface} we presented signal vectors
as a set of types. In order to be completely principled, we should isolate these
types into their own {\em kind} (a sort of type of types); however, the Haskell
extension for this was far from stable at the time this system was created.
\rn{This should probably be a footnote... doesn't need to be highlighted.}

Using the Glasgow Haskell Compiler's extension for Generalized Algebraic
Datatypes~\cite{Cheney2003,Xi2003,PeytonJones2006}, we specify three concrete
datatypes which are parameterized over signal vectors, and represent
information about \textred{one} temporal point \rn{and not others?
  rephrase?} in a signal function's evaluation.

@@

The next representation contains a constructor for {\tt SVEvent} nodes, carrying
a value typed by the type parameter of {\tt SVEvent}, and two constructors, each
of which carries a representation for one child of a {\tt SVAppend} node (the
left or right), and is parametric in the other. \rn{?? Example?} Thus, an instance of this
representation is a path from the root of a signal vector to an event node,
carrying exactly one value for the event.

@@

This is the \textred{type and} framework for our signal function implementation.
Combinators are implemented as functions which return these types with specific
implementations of the continuations. Space precludes a full discussion of the
implemenation of each combinator, but some of the examples will be discussed.

@@

Time dependence is introduced by the {\tt time}, {\tt delay}, and {\tt integrate}
combinators. The time combinator simply sums the time updates and provides the
sum as a signal output. The {\tt delay} combinator keeps a table of events
which have come in, along with their schedule occurrence time, and produces
them as output when time advances far enough. The integrate combinator performs
rectangle-rule integration on signal samples with respect to
time. \rn{repetitive with previous section...}
